---
title: "favar_sdfm"
author: "laura montaldo"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup1, include=FALSE}
#knitr::opts_knit$set(root.dir = '~/Desktop/Tesis_Maestria')
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r}
library(readxl)
library(readr)
library(boot)
library(tsDyn)
library(vars)
library(repr)
library(dplyr)
library(dfms)
library(xts)
library(vars)
library(fbi)
library(forecast)
library(OOS)
library(zoo)
```

```{r}
cat("My Working directory is: ", getwd(), "\n")
df <- read_csv("../../data/prepro/sfr_train.csv", show_col_types = FALSE)
df_test <- read_csv("../../data/prepro/sfr_test.csv", show_col_types = FALSE)
slow <- read_csv("../../data/prepro/slow_columns.csv", show_col_types = FALSE)
fast <- read_csv("../../data/prepro/fast_columns.csv", show_col_types = FALSE)
descr <- read.table("../../data/prepro/descripciones.txt", header = TRUE, sep = "\t")
```

```{r}
data_bal <- as.data.frame(df)
dim(data_bal)
```

```{r}
head(data_bal)
```

```{r}
data_scaled = scale(data_bal[,2:ncol(data_bal)], center = TRUE, scale = TRUE)
data_s=as.data.frame(data_scaled)
rownames(data_s) <- data_bal[,1]
head(data_s)
```


```{r}
dim(data_s)
```
## Paso 1: se extraen los componentes principales de $X_t$ incluyendo $Y_t$ y se determina la cantidad óptima de factores


```{r}
ics = ICr(data_s)
ic_p2_factors=ics$r.star[2]
print(ics)
```



```{r}
plot(ics)
```



```{r}
screeplot(ics, main="Scree plot")
```


```{r}
screeplot(ics, type="ev", main="Scree plot")
```


```{r}
screeplot(ics,type="cum.pve", main="Scree plot")
```


```{r}
# components all
C<-ics$F_pca[, 1:ic_p2_factors]
cat("Dimensiones de los factores comunes en filas y columnas:", dim(C))
```


## Step 2: Extract principal componentes of Slow Variables

```{r}
slow_vars <- unlist(slow$slow)
data_slow <- data_s[, slow_vars]
```

```{r}
dim(data_slow)
```


```{r}
ics_slow = ICr(data_slow)
ic_p2_slow_factors=ics_slow$r.star[2]
```

```{r}
print(ics_slow)
```



```{r}
plot(ics_slow)
screeplot(ics_slow)
screeplot(ics_slow, type="ev")
screeplot(ics_slow,type="cum.pve")
```



```{r}
F_slow<-ics$F_pca[,1:ic_p2_slow_factors]
dim(F_slow)
```
```{r}
dim(C)
```
## Step 3: Clean the PC from the effect of observed Y

```{r}
fedfunds <- as.matrix(data_s[, "FEDFUNDS"])
reg <- lm(C ~ F_slow + fedfunds)
```

### $\hat{F}_t$
```{r}
F_hat <- C - data.matrix(data_s[, "FEDFUNDS"]) %*% reg$coefficients[nrow(reg$coefficients),]
dim(F_hat)
```
## Step 4: Estimate FAVAR and get IRFs

```{r}
data_var <- data.frame(F_hat, "FEDFUNDS" = data_s[, "FEDFUNDS"])
dim(data_var)# common components ?
```

```{r}
var_select <- VARselect(data_var, lag.max = 15, type = "both")
```


```{r}
best_lag <- var_select$selection
best_lag[1] 
```

```{r}
cat("Best # lags according to AIC:", best_lag[1], "\n")
```





```{r}
var = VAR(data_var, p =best_lag[1])
irf_point = irf(var, n.ahead = 48, impulse = "FEDFUNDS", response = "FEDFUNDS", boot = FALSE)
# Shock size of 25 basis points
impulse_sd = 0.25/sd(df$FEDFUNDS)
scale = impulse_sd/(irf_point$irf$FEDFUNDS[1]) # position of FYFF response at step 0
# Computing Loading Factors
matriz_s<- as.matrix(data_s)
matriz_fhat<- as.matrix(F_hat)
reg_loadings = lm(matriz_s ~ 0 + matriz_fhat + data_s[,"FEDFUNDS"])
loadings = reg_loadings$coefficients
#head(reg_loadings$coefficients)
#summary(reg_loadings)
```



```{r}
#### BOOTSTRAPING ########
R = 500 # Number of simulations
nvars = dim(data_s)[2] # Number of variables
nsteps = 49 # numbers of steps
IRFs = array(c(0,0,0), dim = c(nsteps,nvars,R))
var = lineVar(data_var, lag = best_lag[1], include = "const")
for(j in 1:R){    
    data_boot = VAR.boot(var, boot.scheme ="resample")
    var_boot = VAR(data_boot, lag = best_lag[1])
    irf1 = irf(var_boot, n.ahead = 48, impulse = "FEDFUNDS", boot = FALSE)
    for(i in 1:nvars){
        IRFs[,i,j] = (irf1$irf$FEDFUNDS %*% matrix(loadings[, i]))*scale
        }
} ## Boot simulations done

# Extract the quantiles of IRFs we are interested: 90% confidence intervals in BBE
Upper = array(c(0,0), dim = c(nsteps, nvars))
for(k in 1:nsteps){
    for(i in 1:nvars){
        Upper[k,i] = quantile(IRFs[k,i,], probs = c(0.95))[1]
        }
}
Lower = array(c(0,0), dim = c(nsteps, nvars))
for(k in 1:nsteps){
    for(i in 1:nvars){
        Lower[k,i] = quantile(IRFs[k,i,], probs = c(0.05))[1]
        }
}
IRF = array(c(0,0), dim = c(nsteps, nvars))
for(k in 1:nsteps){
    for(i in 1:nvars){
        IRF[k,i] = quantile(IRFs[k,i,], probs = c(0.5))[1]
        }
}
rm(var_boot)
rm(IRFs)
```




```{r}
matching_names <- intersect(names(data_s), descr$fred)
filtered_descr <- descr %>%filter(fred %in% matching_names)
filtered_descr
```




```{r}
variables = c(grep("^FEDFUNDS$", colnames(data_s)),#Fed Funds Rate
              grep("^INDPRO$", colnames(data_s)), #IP Index
              grep("^CPIAUCSL$", colnames(data_s)), #CPI : All Items
            grep("^TB3MS$", colnames(data_s)), #3-Month Treasury Bill:
            grep("^GS5$", colnames(data_s)), #5-Year Treasury Rate
            grep("^M1SL$", colnames(data_s)),# M1 Money Stock
            grep("^M2SL$", colnames(data_s)), #M2 Money Stock
            grep("^EXJPUSx$", colnames(data_s)), #Japan / U.S. Foreign Exchange Rate
            grep("^CUSR0000SAC$", colnames(data_s)),#CPI : Commodities
            grep("^CUMFNS$", colnames(data_s)),#Capacity Utilization: Manufacturing
            grep("^DPCERA3M086SBEA$", colnames(data_s)),#Real personal consumption expenditures
            grep("^DDURRG3M086SBEA$", colnames(data_s)), #Personal Cons. Exp: Durable goods
            grep("^DNDGRG3M086SBEA$", colnames(data_s)), #Personal Cons. Exp: Nondurable goods
            grep("^UNRATE$", colnames(data_s)), #Civilian Unemployment Rate
            grep("^CE16OV$", colnames(data_s)), #Civilian Employment
            grep("^CES0600000008$", colnames(data_s)), #Avg Hourly Earnings : Goods-Producing
            grep("^HOUST$", colnames(data_s)),	#Housing Starts: Total New Privately Owned
            grep("^AMDMNOx$", colnames(data_s)), #New Orders for Durable Goods
            grep("^S&P div yield$", colnames(data_s))#S&P s Composite Common Stock: Dividend Yield
             )
length(variables)
```


```{r}
variable_names = c("Fed Funds Rate", 
"IP Index",
"CPI",
"3-Month Treasury Bill",
"5-Year Treasury Rate",
"M1 Money Stock",
"M2 Money Stock",
"Exchange rate: Japan/U.S.",
"CPI : Commodities",
"Capacity Utilization",
"Real personal consump.",
"Durable goods consump.",
"Nondurable goods consump.",
"Unemployment Rate",
"Employment",
"Avg Hourly Earnings",
"Housing Starts",
"New Orders",
"Dividend Yield")
length(variable_names)
```

```{r}
transf_code = c(1,5,5,1,1,5,5,5,1,1,5,5,5,1,1,5,1,1,1)
length(transf_code)
```


```{r}
options(repr.plot.width=12, repr.plot.height=8)

par(mfrow=c(5,4), 
   mar = c(2, 2, 2, 2))
for(i in variables){
    index = which(variables == i)
    if(transf_code[index] == 5){
        plot(cumsum(IRF[,i]), type ='l',lwd=2, main = variable_names[index],
             ylab= "", xlab="Steps", ylim=range(cumsum(Lower[,i]),cumsum(Upper[,i])),
             cex.main=1, cex.axis=1)
        lines(cumsum(Upper[,i]), lty=2, col="red")
        lines(cumsum(Lower[,i]), lty=2, col="red")
        abline(h=0)
    }
    else{
        plot(IRF[,i], type ='l',lwd=2, main = variable_names[index],
             ylab= "", xlab="Steps", ylim=range((Lower[,i]),(Upper[,i])),
             cex.main=1, cex.axis=1)
        lines((Upper[,i]), lty=2, col="red")
        lines((Lower[,i]), lty=2, col="red")
        abline(h=0)  
    }
}
```



### How much of the variation is captured by the Factors and FFR?


```{r}
options(repr.plot.width=12, repr.plot.height=6)
par(mfrow=c(1, 3), mar = c(2, 2, 2, 2))

plot(data_s$INDPRO, type = "l", ylab = "", main = "Industrial Production", cex.main=1)
lines(fitted(reg_loadings)[,variables[2]], lty=2, col="green")
legend(180,-4, legend=c("Data", "Predicted by PC"), lty = 1:2, col = c("black", "green"), 
      box.lty=0, cex=0.9)
plot(data_s$DPCERA3M086SBEA, type = "l", ylab = "", main = "CPI", cex.main=1)
lines(fitted(reg_loadings)[,variables[3]], lty=2, col="green")
plot(data_s$TB3MS, type = "l", ylab = "", main = "3-month treasury bill", cex.main=1)
lines(fitted(reg_loadings)[,variables[4]], lty=2, col="green")
```


### Regresión de cada factor sobre cada serie




```{r}
# Initialize an empty dataframe to store the results
results_df <- data.frame(Variable = character(0), Factor = character(0), R2 = numeric(0))

# Iterate through each factor in F_hat
for (factor_name in colnames(F_hat)) {
  # Extract the current factor
  current_factor <- F_hat[, factor_name]
  
  # Iterate through each column in data_s
  for (column_name in colnames(data_s)) {
    # Extract the current column
    current_column <- data_s[, column_name]
    
    # Perform linear regression
    regression_model <- lm(current_column ~ current_factor)
    
    # Calculate R-squared
    r_squared <- summary(regression_model)$r.squared
    
    # Add the results to the results dataframe
    results_df <- rbind(results_df, data.frame(Variable = column_name, Factor = factor_name, R2 = r_squared))
  }
}
results_df$R2 <- format(results_df$R2, scientific = FALSE)
results_df <- results_df[order(-as.numeric(results_df$R2)), ]

# Print the sorted results dataframe
print(head(results_df,20))
```

```{r}
average_r_squared_by_factor <- aggregate(R2 ~ Factor, data = results_df, FUN = function(x) mean(as.numeric(x)))

# Print the average R-squared values by factor
print(average_r_squared_by_factor)
```



### Step 5: FEVD


```{r}
# Get the VAR point estimates
hor=60
irf_points = irf(var, n.ahead = hor,  boot = FALSE)
```


```{r}
# Get IRFs for all of X we are interested in, Dimensions: (hor, key_nvars)
# Find loadings
results = summary(reg_loadings) # the warning comes because of FEDFUNDS
key_nvars = length(variables)
key_nvars
```



```{r}
irf_X_pc1 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc2 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc3 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc4 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc5 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc6 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_pc7 = array(c(0,0), dim=c(hor+1, key_nvars))
irf_X_ffr = array(c(0,0), dim=c(hor+1, key_nvars))
for(i in 1:key_nvars){
    irf_X_pc1[,i] = irf_points$irf$PC1 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc2[,i] = irf_points$irf$PC2 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc3[,i] = irf_points$irf$PC3 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc4[,i] = irf_points$irf$PC3 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc5[,i] = irf_points$irf$PC3 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc6[,i] = irf_points$irf$PC3 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_pc7[,i] = irf_points$irf$PC3 %*% matrix(loadings[1:nrow(loadings), variables[i]])
    irf_X_ffr[,i] = (irf_points$irf$FEDFUNDS) %*% matrix(loadings[1:nrow(loadings), variables[i]])
    }
```



```{r}
# Get the IRFs squared and accumulate them
psi2_pc1 = array(0, dim = key_nvars)
psi2_pc2 = array(0, dim = key_nvars)
psi2_pc3 = array(0, dim = key_nvars)
psi2_pc4 = array(0, dim = key_nvars)
psi2_pc5 = array(0, dim = key_nvars)
psi2_pc6 = array(0, dim = key_nvars)
psi2_pc7 = array(0, dim = key_nvars)
psi2_ffr = array(0, dim = key_nvars)

for(i in 1:key_nvars){
    for(j in 1:hor){
        psi2_pc1[i] = psi2_pc1[i] + irf_X_pc1[j,i]^2
        psi2_pc2[i] = psi2_pc2[i] + irf_X_pc2[j,i]^2
        psi2_pc3[i] = psi2_pc3[i] + irf_X_pc3[j,i]^2
        psi2_pc4[i] = psi2_pc4[i] + irf_X_pc4[j,i]^2
        psi2_pc5[i] = psi2_pc5[i] + irf_X_pc5[j,i]^2
        psi2_pc6[i] = psi2_pc6[i] + irf_X_pc6[j,i]^2
        psi2_pc7[i] = psi2_pc7[i] + irf_X_pc7[j,i]^2
        psi2_ffr[i] = psi2_ffr[i] + irf_X_ffr[j,i]^2    
    }
}
```




```{r}
var_total= array(0, dim = key_nvars)
var_fac= array(0, dim = key_nvars)
var_e= array(0, dim = key_nvars)

for(i in 1:key_nvars){
    var_fac[i] = psi2_pc1[i] + psi2_pc2[i] + psi2_pc3[i] + psi2_pc4[i]+psi2_pc5[i]+psi2_pc6[i]+psi2_pc7[i]+psi2_ffr[i]
    var_total[i] = psi2_pc1[i] + psi2_pc2[i] + psi2_pc3[i] +psi2_pc4[i]+psi2_pc5[i]+psi2_pc6[i]+psi2_pc7[i]+ psi2_ffr[i] + results[[variables[i]]]$sigma^2
    var_e[i] = results[[variables[i]]]$sigma^2
}
```





```{r}
table = data.frame("PC1" = round((psi2_pc1),3), 
                   "PC2" = round((psi2_pc2),3), 
                   "PC3" = round((psi2_pc3),3), 
                   "PC4" = round((psi2_pc4),3),
                   "PC5" = round((psi2_pc5),3),
                   "PC6" = round((psi2_pc6),3),
                   "PC7" = round((psi2_pc7),3),
                   "FEDFUNDS" = round((psi2_ffr),3),
                  "Factor_Y_total" = round(var_fac,3) ,
                  "e" = round((var_e),3),
                  "Total" = round(var_total,3))
row.names(table) = variable_names
table
```

```{r}
r2 = array(0, dim = key_nvars)
for(i in 1:key_nvars){
    r2[i] = results[[variables[i]]]$r.squared
}

table2 = data.frame("Variables" = variable_names, "Contribution" = round((psi2_ffr/var_total),3), "R-squared" = round(r2,3))
table2
```


## Model Performance



```{r}
data_test <- as.data.frame(df_test)
test_scaled = scale(data_test[,2:ncol(data_test)], center = TRUE, scale = TRUE)
test_s=as.data.frame(test_scaled)
rownames(test_s) <- data_test[,1]
head(test_s)
```

### OOS

```{r}
dim(matriz_s)
```



```{r}
INDPRO=data.frame(data_s[,"INDPRO"])
FEDFUNDS=data.frame(data_s[,"FEDFUNDS"])
colnames(INDPRO) <- "INDPRO"
colnames(FEDFUNDS) <- "FEDFUNDS"
x_f=cbind(INDPRO,FEDFUNDS,data.frame(matriz_fhat))
date=data_bal[,1]
Data = data.frame(date = date, x_f )
names(Data)
```




```{r}
start_date <- as.Date("2007-08-01")
# Add 136 months to the start date
end_date <- as.Date("2020-12-01")
# Calculate the difference in months
diff_months <- as.numeric(difftime(end_date, start_date, units = "days") / 30.44)
# Convert the result to an integer
diff_months <- round(diff_months)
# Print the result
cat("Number of months between the two test dates:", diff_months, "\n")
```

```{r}
forecast_multivariate(Data, target= "INDPRO", forecast.dates = tail(Data$date,diff_months+1),method='ols',horizon = diff_months+1,freq = 'month', outlier.clean = FALSE,
       impute.missing = FALSE,lag.n=NULL )
```
```{r}
class(colnames(data_s))
```


```{r}
forecast_results_list <- list()
# Define the function to run forecast_multivariate for a given variable name
forecast_function <- function(var_name) {
  forecast_result <- forecast_multivariate(
    Data = Data,
    target = var_name,
    forecast.dates = tail(Data$date, diff_months + 1),
    method = 'ols',
    horizon = diff_months + 1,
    freq = 'month',
    outlier.clean = FALSE,
    impute.missing = FALSE,
    lag.n = NULL
  )
  return(forecast_result)
}

# Apply the forecast_function to each variable name using lapply
forecast_results_list <- lapply(as.list(colnames(data_s)), forecast_function)
```

```{r}
list(names(data_s))[[1]]
```





